---
title: 'Data Brief #1'
author: "Mutaal Tahir"
date: "`r Sys.Date()`"
output: pdf_document
editor_options: 
  markdown: 
    wrap: 72
---

# Research Topic and Question

The topic that I plan on analyzing is the difference in opinions towards
government response and restrictions due to the COVID-19 pandemic among
people from rural and urban Alberta. To address this topic, I plan on
analyzing the September 2021 Alberta Viewpoint Dataset. The motivation
behind choosing this data set is that it marked over a year of COVID-19
restrictions being in place in Alberta and I aim for finding how they
have felt so far with the government\'s response to the pandemic
particularly from a rural/urban perspective. Do the opinions of rural
residents differ from their counterparts in the urban areas or do they
feel similarly? If they have differences, where do these differences
mainly come from?

We know from prior research that there has been a rural urban divide
present in Alberta (Banack 2021, Epp 2019). This divide has existed due
to various factors and has caused rural residents to feel overlooked,
looked down upon and treated unfairly ​​(Banack, 2021). Epp (2019) argues
that rural sensibility is hard to find in the government caucus, which
reflects on Alberta becoming an increasingly urban province compared to
a rural one. COVID-19 brought more issues that may have fueled this
divide. Among these issues include shining a light on Alberta\'s rural
healthcare system which has already had issues such as the lack of
access for osteoarthritis care despite the high prevalence (Liu et al.,
2022). It was reported by the Canadian Medical Association in 2013 that
21% of the Canadian population lives in rural areas however only 9.4% of
general practitioners and 3% of specialists do (CMA, 2013). COVID-19
causes many infected individuals to be hospitalized which could prove to
be a problem for the already understaffed rural healthcare system.
Torrie et al. (2021) highlights how there have been vaccine distribution
issues in rural Alberta in the past giving an example of the 2009 HINI
influenza pandemic.

I anticipate exploring whether did COVID-19 further increase the divide
where rural residents felt a difference in the government response
compared to the urban residents. How do they find their issues being
handled and whom do they feel is doing a better job?  Do Albertans feel
the federal government is handling the pandemic well or is the
provincial government doing a better job? Furthermore, I set out to
explore how do rural and urban residents personally feel about policies
such as masking or vaccine mandates.

# Data

The 2021 Alberta Viewpoint Survey contains data from 1204 respondents.
Data was collected through a roughly 15-minute voluntary survey. Consent
of respondents was taken before the survey began. Respondents included
individuals from various backgrounds, age, demographics and genders. The
survey had 8 sections namely:

1.  Demographics

2.  Electoral Politics

3.  Campaign Finances and Local Elections

4.  Dialogue and Polarization

5.  Federalism

6.  The COVID-19 Pandemic

7.  Economic Perceptions

8.  Identity and Discrimination

There was missing information in the responses and was deleted using
listwise deletion in order to have clean data to analyze.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r Reading Libraries, include=FALSE}
library(dplyr)
library(ggplot2)
library(descr)
library(DescTools)
library(survey)
```

```{r Display Dimensions}
d1 <- read.csv("ViewpointAB_Sept2021_Survey_Data.csv",header=TRUE)
d2 <- subset(d1,select=c(Q4,Q60r1,Q60r2,Q61r1,Q61r2,Q67,Q68,Q70,Q71))
d3 <- subset(d2,complete.cases(d2),drop=TRUE)
dim(d1)
```

# Key Measures

For ease or reading and simplicity I will divide each variable into it's
own category and describe it.

## Area 

Areas are divided into three categories: urban, suburban and rural.
Urban has been coded into '1', Suburban '2' and rural '3'. I have
recoded them into their respective names so that the viewer does not
have to look back and forth for what each number represents. I also
counted suburban residents as urban residents in this project due to
their close proximity with urban areas.

```{r}
d3$zone=recode(d3$Q4,`1`="Urban",`2`="Urban",`3`="Rural")

round(prop.table(table(d3$zone)) * 100, 2)

ggplot(data = d3, aes(x = zone, y = ..count.. / sum(..count..))) + geom_bar(fill = "darkgreen", color = "black") + scale_y_continuous(labels = scales::percent, limits = c(0,1)) +
labs(x = "Zone",
y = "Percentage",
title = "Participants by Zone") +
theme(axis.text.x = element_text(angle = 45, hjust = 1))


```

Data shows that 87.38% of our respondents are from urban areas whereas
only 12.62% are from rural areas. Since this is a predictor variable and
I base my conclusions on zone area, I will further subset my data into
rural and urban residents. This will allow me to separately measure
responses from both sides.

```{r}
d4 <- subset(d3, zone=="Urban")
d5 <- subset(d3, zone=="Rural")
```

## Trust of Leader (Justin Trudeau the Prime Minister) 

This is an outcome variable I will need to predict how much do rural vs
urban Albertans trust the Prime Minister handling the pandemic. Values
are in numbers and were recoded as: '1' No trust, '2' Little Trust, '3'
A lot of Trust, '4' Full Trust, '98' Not Sure.

```{r Trust of JT by Urban Residents}
d4$trust_jt=recode(d4$Q60r1,`1`="No Trust",`2`="Little Trust",`3`="A lot of Trust",`4`="Full Trust",`98`="Not Sure")
d4$trust_jt <- factor(d4$trust_jt,c("No Trust","Little Trust","A lot of Trust","Full Trust","Not Sure"))
freq(ordered(d4$trust_jt), plot = FALSE)
Mode <- function(x){
  a=table(x)
  return(a[which.max(a)])
}
Mode(d4$trust_jt)

ggplot(data = d4, aes(x = trust_jt, y = ..count.. / sum(..count..))) + geom_bar(fill = "skyblue", color = "black") + scale_y_continuous(labels = scales::percent, limits = c(0,0.4)) +
labs(x = "Trust Level",
y = "Percentage",
title = "Trust Justin Trudeau's Handling of the pandemic by Urban Residents") +
theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

35.84% of Urban residents had to trust on Justin Trudeau's handling of
this pandemic followed by 30.13% having little trust, 21.77% A lot of
trust, 10.17% Full trust and 2.09% not sure.

No trust was the most common response with 377 votes.

```{r Trust of JT by Rural Residents}
d5$trust_jt=recode(d5$Q60r1,`1`="No Trust",`2`="Little Trust",`3`="A lot of Trust",`4`="Full Trust",`98`="Not Sure")
d5$trust_jt <- factor(d5$trust_jt,c("No Trust","Little Trust","A lot of Trust","Full Trust","Not Sure"))
freq(ordered(d5$trust_jt), plot = FALSE)
Mode <- function(x){
  a=table(x)
  return(a[which.max(a)])
}
Mode(d5$trust_jt)

ggplot(data = d5, aes(x = trust_jt, y = ..count.. / sum(..count..))) + geom_bar(fill = "red", color = "black") + scale_y_continuous(labels = scales::percent, limits = c(0,0.6)) +
labs(x = "Trust Level",
y = "Percentage",
title = "Trust Justin Trudeau's Handling of the pandemic by Rural Residents") +
theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

53.29% of Rural residents had to trust on Justin Trudeau's handling of
this pandemic followed by 31.58% having little trust, 11.84% A lot of
trust, 3.29% Full trust. None of the respondents choose "Not sure".

No trust was the most common response with 81 votes.

## Trust of Leader (Jason Kenny then Premier of Alberta) 

This is an outcome variable I will need to predict how much do rural vs
urban Albertans trust the Premier handling the pandemic. Values are in
numbers and were recoded as: '1' No trust, '2' Little Trust, '3' A lot
of Trust, '4' Full Trust, '98' Not Sure.

```{r Trust of JK by Urban Residents}
d4$trust_jk=recode(d4$Q60r2,`1`="No Trust",`2`="Little Trust",`3`="A lot of Trust",`4`="Full Trust",`98`="Not Sure")
d4$trust_jk <- factor(d4$trust_jk,c("No Trust","Little Trust","A lot of Trust","Full Trust","Not Sure"))
freq(ordered(d4$trust_jk), plot = FALSE)
Mode <- function(x){
  a=table(x)
  return(a[which.max(a)])
}
Mode(d4$trust_jk)

ggplot(data = d4, aes(x = trust_jk, y = ..count.. / sum(..count..))) + geom_bar(fill = "skyblue", color = "black") + scale_y_continuous(labels = scales::percent, limits = c(0,0.7)) +
labs(x = "Trust Level",
y = "Percentage",
title = "Trust Jason Kenny's Handling of the pandemic by Urban Residents") +
theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

63.40% of Urban residents had to trust on Jason Kenny's handling of this
pandemic followed by 22.24% having little trust, 9.03% A lot of trust,
2.76% Full trust and 2.57% not sure.

No trust was the most common response with 667 votes.

```{r Trust of JK by Rural Residents}
d5$trust_jk=recode(d5$Q60r2,`1`="No Trust",`2`="Little Trust",`3`="A lot of Trust",`4`="Full Trust",`98`="Not Sure")
d5$trust_jk <- factor(d5$trust_jk,c("No Trust","Little Trust","A lot of Trust","Full Trust","Not Sure"))
freq(ordered(d5$trust_jk), plot = FALSE)
Mode <- function(x){
  a=table(x)
  return(a[which.max(a)])
}
Mode(d5$trust_jk)

ggplot(data = d5, aes(x = trust_jk, y = ..count.. / sum(..count..))) + geom_bar(fill = "red", color = "black") + scale_y_continuous(labels = scales::percent, limits = c(0,0.7)) +
labs(x = "Trust Level",
y = "Percentage",
title = "Trust Jason Kenny's Handling of the pandemic by Rural Residents") +
theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

61.84% of Rural residents had to trust on Jason Kenny's handling of this
pandemic followed by 28.95% having little trust, 7.89% A lot of trust,
0.65% Full trust and 0.65% not sure.

No trust was the most common response with 94 votes.

## Performance of Government (Federal) 

This is an outcome variable I will need to predict how well do rural vs
urban Albertans think the Federal government handled the pandemic.
Values are in numbers and were recoded as: '1' No trust, '2' Little
Trust, '3' A lot of Trust, '4' Full Trust, '98' Not Sure.

```{r Performance (Federal) Rating by Urban Residents}
d4$trust_fg=recode(d4$Q61r1,`1`="Very bad",`2`="Bad",`3`="Good",`4`="Very Good")
d4$trust_fg <- factor(d4$trust_fg,c("Very bad","Bad","Good","Very Good"))
freq(ordered(d4$trust_fg), plot = FALSE)
Mode <- function(x){
  a=table(x)
  return(a[which.max(a)])
}
Mode(d4$trust_fg)

ggplot(data = d4, aes(x = trust_fg, y = ..count.. / sum(..count..))) + geom_bar(fill = "skyblue", color = "black") + scale_y_continuous(labels = scales::percent, limits = c(0,0.7)) +
labs(x = "Performance Rating",
y = "Percentage",
title = "Pandemic Performance Rating of Federal Government by Urban Residents") +
theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

21.48% of urban respondents thought the federal government is doing a
very bad job handling the pandemic. 23.95% thought it was doing a bad
job whereas 46% the federal government was doing a good job making it
the most popular answer with 484 votes. 8.56% thought the government was
doing a very good job.

```{r Performance (Federal) Rating by Rural Residents}
d5$trust_fg=recode(d5$Q61r1,`1`="Very bad",`2`="Bad",`3`="Good",`4`="Very Good")
d5$trust_fg <- factor(d5$trust_fg,c("Very bad","Bad","Good","Very Good"))
freq(ordered(d5$trust_fg), plot = FALSE)
Mode <- function(x){
  a=table(x)
  return(a[which.max(a)])
}
Mode(d5$trust_fg)

ggplot(data = d5, aes(x = trust_fg, y = ..count.. / sum(..count..))) + geom_bar(fill = "red", color = "black") + scale_y_continuous(labels = scales::percent, limits = c(0,0.4)) +
labs(x = "Performance Rating",
y = "Percentage",
title = "Pandemic Performance Rating of Federal Government by Rural Residents") +
theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

31.58% of rural respondents thought the federal government is doing a
very bad job handling the pandemic. 28.28% thought it was doing a bad
job whereas 38.2% the federal government was doing a good job making it
the most popular answer with 58 votes. 1.97% thought the government was
doing a very good job.

## Performance of Government (Provincial) 

This is an outcome variable I will need to predict how well do rural vs
urban Albertans think the provincial government handled the pandemic.
Values are in numbers and were recoded as: '1' No trust, '2' Little
Trust, '3' A lot of Trust, '4' Full Trust, '98' Not Sure.

```{r Performance (Provincial) Rating by Urban Residents}
d4$trust_pg=recode(d4$Q61r2,`1`="Very bad",`2`="Bad",`3`="Good",`4`="Very Good")
d4$trust_pg <- factor(d4$trust_pg,c("Very bad","Bad","Good","Very Good"))
freq(ordered(d4$trust_pg), plot = FALSE)
Mode <- function(x){
  a=table(x)
  return(a[which.max(a)])
}
Mode(d4$trust_pg)

ggplot(data = d4, aes(x = trust_pg, y = ..count.. / sum(..count..))) + geom_bar(fill = "skyblue", color = "black") + scale_y_continuous(labels = scales::percent, limits = c(0,0.7)) +
labs(x = "Performance Rating",
y = "Percentage",
title = "Pandemic Performance Rating of Provincial Government by Urban Residents") +
theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

54.75% of urban respondents thought the provincial government is doing a
very bad job handling the pandemic making it the most popular answer
with 576 votes. 26.71% thought it was doing a bad job whereas 15.5% the
federal government was doing a good job . 3.0% thought the provincial
government was doing a very good job.

```{r Performance (Provincial) score by Rural Residents}
d5$trust_pg=recode(d5$Q61r2,`1`="Very bad",`2`="Bad",`3`="Good",`4`="Very Good")
d5$trust_pg <- factor(d5$trust_pg,c("Very bad","Bad","Good","Very Good"))
freq(ordered(d5$trust_pg), plot = FALSE)
Mode <- function(x){
  a=table(x)
  return(a[which.max(a)])
}
Mode(d5$trust_pg)

ggplot(data = d5, aes(x = trust_pg, y = ..count.. / sum(..count..))) + geom_bar(fill = "red", color = "black") + scale_y_continuous(labels = scales::percent, limits = c(0,0.7)) +
labs(x = "Performance Rating",
y = "Percentage",
title = "Pandemic Performance Rating of Provincial Government by Rural Residents") +
theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

50.66% of urban respondents thought the provincial government is doing a
very bad job handling the pandemic making it the most popular answer
with 77 votes. 32.23% thought it was doing a bad job whereas 16.45% the
federal government was doing a good job . 0.66% thought the provincial
government was doing a very good job.

## Support of businesses requiring proof of vaccination (Q67) 

This is an outcome variable I will need to predict how well do rural vs
urban Albertans support the government law allowing businesses to stay
open provided they require proof of vaccination of negative COVID-19
test result . Values are in numbers and were recoded as: '1' Strongly
support, '2' Somewhat support, '3' Neither support or oppose, '4'
Somewhat oppose, '5' Strongly oppose.

```{r Level of support (Business Vaccine) Urban}
d4$business_vac=recode(d4$Q67,`1`="Strongly support",`2`="Somewhat support",`3`="Neither support or oppose",`4`="Somewhat oppose", `5`="Strongly oppose")
d4$business_vac <- factor(d4$business_vac,c("Strongly support","Somewhat support","Neither support or oppose","Somewhat oppose","Strongly oppose"))
freq(ordered(d4$business_vac), plot = FALSE)
Mode <- function(x){
  a=table(x)
  return(a[which.max(a)])
}
Mode(d4$business_vac)

ggplot(data = d4, aes(x = business_vac, y = ..count.. / sum(..count..))) + geom_bar(fill = "skyblue", color = "black") + scale_y_continuous(labels = scales::percent, limits = c(0,0.7)) +
labs(x = "Support Level",
y = "Percentage",
title = "Vaccine Proof Support for Businesses in Urban Alberta") +
theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

49.72% of urban respondents strongly supported businesses requiring
proof of vaccination or negative COVID-19 tests making it the most
popular response with 523 votes. 23.57% supported the statement,
followed by 9.8% being neutral, 4.28% opposing and 12.64% strongly
opposing the statement.

```{r Level of support (Business Vaccine) Rural}
d5$business_vac=recode(d5$Q67,`1`="Strongly support",`2`="Somewhat support",`3`="Neither support or oppose",`4`="Somewhat oppose", `5`="Strongly oppose")
d5$business_vac <- factor(d5$business_vac,c("Strongly support","Somewhat support","Neither support or oppose","Somewhat oppose","Strongly oppose"))
freq(ordered(d5$business_vac), plot = FALSE)
Mode <- function(x){
  a=table(x)
  return(a[which.max(a)])
}
Mode(d5$business_vac)

ggplot(data = d5, aes(x = business_vac, y = ..count.. / sum(..count..))) + geom_bar(fill = "red", color = "black") + scale_y_continuous(labels = scales::percent, limits = c(0,0.7)) +
labs(x = "Support Level",
y = "Percentage",
title = "Vaccine Proof Support for Businesses in Rural Alberta") +
theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

46.1% of rural respondents strongly supported businesses requiring proof
of vaccination or negative COVID-19 tests making it the most popular
response with 70 votes. 18.42% supported the statement, followed by
9.87% being neutral, 5.92% opposing and 19.74% strongly opposing the
statement.

## Support for vaccine passports (Q68) 

This is an outcome variable I will need to predict how well do rural vs
urban Albertans support vaccine passports. Values are in numbers and
were recoded as: '1' Strongly support, '2' Somewhat support, '3' Neither
support or oppose, '4' Somewhat oppose, '5' Strongly oppose.

```{r Level of support (Vaccine pass) Urban}
d4$vac_pass=recode(d4$Q68,`1`="Strongly support",`2`="Somewhat support",`3`="Neither support or oppose",`4`="Somewhat oppose", `5`="Strongly oppose")
d4$vac_pass <- factor(d4$vac_pass,c("Strongly support","Somewhat support","Neither support or oppose","Somewhat oppose","Strongly oppose"))
freq(ordered(d4$vac_pass), plot = FALSE)
Mode <- function(x){
  a=table(x)
  return(a[which.max(a)])
}
Mode(d4$vac_pass)

ggplot(data = d4, aes(x = vac_pass, y = ..count.. / sum(..count..))) + geom_bar(fill = "skyblue", color = "black") + scale_y_continuous(labels = scales::percent, limits = c(0,0.7)) +
labs(x = "Support Level",
y = "Percentage",
title = "Support for Vaccine Passport Urban Alberta") +
theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

62.74% of urban residents supported vaccine passports making it the most
popular option with 660 votes. This was followed by 14.45% somewhat
supporting, 7.32% neutral, 2.95% opposing and 12.55% strongly opposing.

```{r Level of support (Vaccine pass) Rural}
d5$vac_pass=recode(d5$Q68,`1`="Strongly support",`2`="Somewhat support",`3`="Neither support or oppose",`4`="Somewhat oppose", `5`="Strongly oppose")
d5$vac_pass <- factor(d5$vac_pass,c("Strongly support","Somewhat support","Neither support or oppose","Somewhat oppose","Strongly oppose"))
freq(ordered(d5$vac_pass), plot = FALSE)
Mode <- function(x){
  a=table(x)
  return(a[which.max(a)])
}
Mode(d5$vac_pass)

ggplot(data = d5, aes(x = vac_pass, y = ..count.. / sum(..count..))) + geom_bar(fill = "red", color = "black") + scale_y_continuous(labels = scales::percent, limits = c(0,0.7)) +
labs(x = "Support Level",
y = "Percentage",
title = "Support for Vaccine Passport Rural Alberta") +
theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

55.26% of rural residents supported vaccine passports making it the most
popular option with 84 votes. This was followed by 9.87% somewhat
supporting, 7.23% neutral, 5.26% opposing and 22.37% strongly opposing.

## Feelings on current restrictions (Q70) 

This is an outcome variable I will need to predict how well do rural vs
urban Albertans feel about current restrictions. Values are in numbers
and were recoded as: '1' Too harsh, '2' Too lenient, '3' Just about
right.

```{r Level of support (Current Restrictions) Urban}
d4$feel_cr=recode(d4$Q70,`1`="Too harsh",`2`="Too lenient",`3`="Just about right")
d4$feel_cr <- factor(d4$feel_cr,c("Too harsh","Too lenient","Just about right"))
freq(ordered(d4$feel_cr), plot = FALSE)
Mode <- function(x){
  a=table(x)
  return(a[which.max(a)])
}
Mode(d4$feel_cr)

ggplot(data = d4, aes(x = feel_cr, y = ..count.. / sum(..count..))) + geom_bar(fill = "skyblue", color = "black") + scale_y_continuous(labels = scales::percent, limits = c(0,0.7)) +
labs(x = "Feelings",
y = "Percentage",
title = "Feelings on Current Restrictions Urban Alberta") +
theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

14.83% of Urban residents found current restrictions to be too harsh
followed by 57.6% saying they are too lenient making it the most common
response with 606 votes. 27.57% found them to be just about right.

```{r Level of support (Current Restrictions) Rural}
d5$feel_cr=recode(d5$Q70,`1`="Too harsh",`2`="Too lenient",`3`="Just about right")
d5$feel_cr <- factor(d5$feel_cr,c("Too harsh","Too lenient","Just about right"))
freq(ordered(d5$feel_cr), plot = FALSE)
Mode <- function(x){
  a=table(x)
  return(a[which.max(a)])
}
Mode(d5$feel_cr)

ggplot(data = d5, aes(x = feel_cr, y = ..count.. / sum(..count..))) + geom_bar(fill = "red", color = "black") + scale_y_continuous(labels = scales::percent, limits = c(0,0.7)) +
labs(x = "Feelings",
y = "Percentage",
title = "Feelings on Current Restrictions Rural Alberta") +
theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

23.03% of rural residents found current restrictions to be too harsh
followed by 53.95% saying they are too lenient making it the most common
response with 82 votes. 23.03% found them to be just about right.

## Masks: Freedom or public health (Q71) 

This is an outcome variable I will need to predict how do rural vs urban
Albertans feel about masks. Does freedom come before public health or
viceversa? Values are in numbers and were recoded as: '1' Freedom, '2'
Public Health.

```{r Level of support (Masks) Urban}
d4$masks=recode(d4$Q71,`1`="Freedom",`2`="Public Health")
d4$masks <- factor(d4$masks,c("Freedom","Public Health"))
freq(ordered(d4$masks), plot = FALSE)
Mode <- function(x){
  a=table(x)
  return(a[which.max(a)])
}
Mode(d4$masks)

ggplot(data = d4, aes(x = masks, y = ..count.. / sum(..count..))) + geom_bar(fill = "skyblue", color = "black") + scale_y_continuous(labels = scales::percent, limits = c(0,0.9)) +
labs(x = "Freedom vs Public Health",
y = "Percentage",
title = "Masks: Freedom or Public Health Urban Alberta") +
theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

14.35% of Urban residents choose freedom over public health when
discussing about face masks while 85.65% choose public health making it
the most common response with 901 votes.

```{r Level of support (Masks) Rural}
d5$masks=recode(d5$Q71,`1`="Freedom",`2`="Public Health")
d5$masks <- factor(d5$masks,c("Freedom","Public Health"))
freq(ordered(d5$masks), plot = FALSE)
Mode <- function(x){
  a=table(x)
  return(a[which.max(a)])
}
Mode(d5$masks)

ggplot(data = d5, aes(x = masks, y = ..count.. / sum(..count..))) + geom_bar(fill = "red", color = "black") + scale_y_continuous(labels = scales::percent, limits = c(0,0.9)) +
labs(x = "Freedom vs Public Health",
y = "Percentage",
title = "Masks: Freedom or Public Health Rural Albert") +
theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

18.42% of rural residents choose freedom over public health when
discussing about face masks while 81.58% choose public health making it
the most common response with 124 votes.

## References

Banack, C. (2020, October 26). Ethnography and Political Opinion:
Identity, Alienation  and Anti-establishmentarianism in Rural Alberta.
Canadian Journal of Political Science, 54(1), 1--22.
<https://doi.org/10.1017/s0008423920000694>

Epp, Roger. 2019. \"The End of Exceptionalism: Post-rural Politics in
Alberta.\" In Orange Chinook: Politics in the New Alberta, ed. Duane
Bratt, Keith Brownsey, Richard Sutherland and David Taras. Calgary:
University of Calgary Press.

Liu, X.; Seidel, J.E.; McDonald, T.; Patel, A.B.; Waters, N.; Bertazzon,
S.; Shahid, R.; Marshall, D.A. Rural--Urban Disparities in Realized
Spatial Access to General Practitioners, Orthopedic Surgeons, and
Physiotherapists among People with Osteoarthritis in Alberta, Canada.
Int. J. Environ. Res. Public Health**2022**,19,7706. https://
doi.org/10.3390/ijerph19137706

Canadian Medical Association. CMA Position Statement: Ensuring Equitable
Access to Care: Strategies for Governments, Health System Planners, and
the Medical Profession; Canadian Medical Association: Ottawa, ON,
Canada, 2013.

Torrie, C., Yanicki, S., Sedgwick, M., & Howard, L. (2021, February 1).
Social justice in pandemic
immunization policy: We\'re all in this together. *Nursing Ethics*,
*28*(6),
924--934. <https://doi.org/10.1177/0969733020983395>
